\chapter{Introduction}

This blueprint presents the formalized universal approximation theorems (UATs) for neural networks
developed in Lean~4 and Mathlib, across six developments. Each result is stated informally,
accompanied by its formal Lean statement and a prose proof, and is cross-linked to the formal API.
Every item marked green is fully formalized and \texttt{sorry}-free.

\begin{itemize}
  \item \textbf{Cybenko (1989)}: a single-hidden-layer network with a continuous sigmoidal
    activation is dense in $\Ck$.
  \item \textbf{Leshno--Lin--Pinkus--Schocken (1993)}: an $M$-class activation densely
    approximates iff it is not (a.e.) a polynomial.
  \item \textbf{Mikulincer--Reichman (2022)}: depth-4 monotone threshold networks exactly
    interpolate and uniformly approximate any monotone function.
  \item \textbf{Sartor et al.\ (2025)}: monotone one-sided-saturating activations are universal,
    extending beyond bounded activations.
  \item \textbf{Runje et al.\ (forthcoming)}: \emph{Deep Constrained Monotonic Neural Networks} ---
    skip connections make deep constrained monotone networks trainable, with formalized soundness
    (a residual stack of any depth is monotone) and universal approximation; partial monotonicity
    is a secondary result. Extends Runje--Shankaranarayana 2023.
  \item \textbf{Amos--Xu--Kolter (2017)}: \emph{Input-Convex Neural Networks} --- fully
    input-convex networks (FICNNs) denote convex functions of their input; ICNN soundness is
    formalized, with the convex universal approximation property forthcoming.
\end{itemize}

The three monotone developments (Mikulincer--Reichman, Sartor, Runje) share a common,
activation-generic \texttt{ActStack} core.
